---
title: "Priority Neighbourhoods in Aberdeen"

output:
  flexdashboard::flex_dashboard:
    vertical_layout: scroll 
    orientation: rows
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(janitor)
library(sf)
library(flexdashboard)
library(mapview)
library(gt)
library(kableExtra)
library(RColorBrewer)
```

```{r}
#Load data

#SIMD and census data combined
#This file was created using data_processing.Rmd
scotland_measures <-
  read_csv("scotland_deprivation_with_rankings.csv")

#2011 census datazones for maps
datazone_sf <- 
  st_read(here(
    "data", 
    "sc_dz_11.shp"), quiet = T) %>%
  clean_names()

#Priority areas
#source 
#https://communityplanningaberdeen.org.uk/priority-neighbourhoods/

priority_areas <-
  c("Torry", "Middlefield", "Mastrick", "Cummings Park", "Northfield", "Heathryfold", "Seaton", "Woodside", "Tillydrone", "Kincorth", "Ashgrove", "Stockethill", "George Street")


#filter Aberdeen Priority Neighbourhood data only
#exclude the South part of Kincorth  

aberdeen_priority <-
  scotland_measures %>%
  filter(la_name == "Aberdeen City") %>%
  mutate(priority_area = 
           str_detect(dz_name, 
                      paste(priority_areas, collapse="|")),
         priority_area =
           ifelse(
             dz_name == 
               "Kincorth, Leggart and Nigg South - 01",
             F, 
             priority_area)) %>%
  filter(priority_area == T)

aberdeen_data_zones <- pull(aberdeen_priority, data_zone)

#Aberdeen shapefile for maps
aberdeen_sf <- 
  filter(datazone_sf, 
         data_zone %in% aberdeen_data_zones) 

aberdeen_priority_sf <-
  merge(aberdeen_sf, aberdeen_priority, by = "data_zone") %>%
  filter(priority_area == T) %>%
  select(data_zone, name, iz_name, simd2020v2_decile, ends_with("scot"), geometry) 
```

Neighbourhoods {data-orientation="columns"}
==============================================

Column
----------------------------------------------

### 

This map shows Aberdeen's [**Priority Neighbourhoods**](https://communityplanningaberdeen.org.uk/priority-neighbourhoods/).

The Community Empowerment Act gives support to neighbourhoods where people have poor outcomes due to socio-economic disadvantage.

This website shows which Priority Neighbourhoods need support for

- central heating
- overcrowded homes
- lone parents
- recent immigrants
- young people not in education, employment, or training
- hospitalisations
- crime

**See the menu above for more information about Priority Neighbourhoods.**

Column
----------------------------------------------

### 

```{r}
aberdeen_priority_sf %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "iz_name", label = "name",
    layer.name = "Priority Neighbourhoods",
    color = "black", lwd = 1)
```

Deprivation
==============================================

Row
----------------------------------------------

### 

The Scottish Government measures disadvantage with a score called the [Scottish Index of Multiple Deprivation](https://www.gov.scot/collections/scottish-index-of-multiple-deprivation-2020/). This score is created using data from 40 measures of health, education, crime, housing, and income.

This table shows **all of the Aberdeen Priority Neighbourhoods with their total disadvantage score**.

Woodside (zone 4) is the most disadvantaged in Aberdeen, with a score of 97. A score of 97 means this Woodside neighbourhood is more disadvantaged than 97% of Scotland.

**The dark red and orange areas are the priority neighbourhoods with the most disadvantage overall.**

Row
----------------------------------------------

### 

```{r}
#SIMD
aberdeen_priority_sf %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "simd_rank_scot",
    label = "name",
    layer.name = "Deprivation<br>Percentile<br>in Scotland",
    color = "black", lwd = 1, 
    alpha.regions = 1, 
    col.regions = rev(brewer.pal(100, name = "Spectral")))
```

### 

```{r}
#areas by deprivation

aberdeen_priority %>%
  select(data_zone, dz_name, total_population, simd_rank_scot) %>%
  arrange(desc(simd_rank_scot)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by most disadvantaged overall") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    simd_rank_scot = "deprivation percentile in Scotland")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Housing
==============================================

Row
----------------------------------------------

### 

Here are **two measures of housing quality in Priority Neighbourhoods**

-   How many people have no central heating?

-   How many people live in overcrowded homes?

Row
----------------------------------------------

### 

**Red areas have more people living without central heating than 95% of Scotland.**

```{r}
#central heating
aberdeen_priority_sf %>%
  filter(no_central_heat_rank_scot < 5) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}
aberdeen_priority %>%
  select(data_zone, dz_name, total_population, nocentralheat_rate) %>%
  arrange(desc(nocentralheat_rate)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by most living without central heating") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    nocentralheat_rate = "% no central heat")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Row
----------------------------------------------

### 

**Red areas have more people living in overcrowded homes than 95% of Scotland.**

```{r}
#overcrowding
aberdeen_priority_sf %>%
  filter(overcrowded_rank_scot <= 5) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}
aberdeen_priority %>%
  select(data_zone, dz_name, total_population, overcrowded_rate) %>%
  arrange(desc(overcrowded_rate)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by most living in overcrowded homes") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    overcrowded_rate = "% living overcrowded")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Children
==============================================

Row
----------------------------------------------

### 

Here are **three measures of children's wellbeing in Priority Neighbourhoods**

-   How many people live in lone parent households?

-   How many young people (aged 16 to 19) are not in education, employment, or training?

-   Which areas have the most children under age 5?

Row
----------------------------------------------

### 

**Red areas have more people living in lone parent households than 90% of Scotland.**

```{r}
#lone parents
aberdeen_priority_sf %>%
  filter(lone_parent_rank_scot <= 10) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}
aberdeen_priority %>%
  mutate(lone_parent_perc = lone_parent_prop * 100) %>%
  select(data_zone, dz_name, total_population, lone_parent_perc) %>%
  arrange(desc(lone_parent_perc)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods with most Lone Parents", 
    subtitle = "ordered by most people living in lone-parent households") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    lone_parent_perc = "% living lone parent household")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Row
----------------------------------------------

### 

**Red areas have more young people (aged 16 to 19) not in education, employment, or training than 98% of Scotland.**

```{r}
#neet
aberdeen_priority_sf %>%
  filter(neet_rank_scot <= 2) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}
aberdeen_priority %>%
  select(data_zone, dz_name, total_population, not_participating) %>%
  arrange(desc(not_participating)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by most 16 to 19 year olds not in education, employment, or training") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    not_participating = "% not in education, employment, training")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Row
----------------------------------------------

### 

**Red areas have more children under age 5 than 95% of Scotland.**

```{r}
#under 5s
aberdeen_priority_sf %>%
  filter(under_5_rank_scot <= 5) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}
aberdeen_priority %>%
  mutate(under_5_perc = under_5_prop * 100) %>%
  select(data_zone, dz_name, total_population, under_5_perc) %>%
  arrange(desc(under_5_perc)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by most under age 5") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    under_5_perc = "% under age 5")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Immigration
==============================================

Row
----------------------------------------------

### 

Here are **two measures of immigration and ethnicity in Priority Neighbourhoods**

-   How many people have recently immigrated into the UK?

-   How many people are from minority ethnic backgrounds?

Row
----------------------------------------------

### 

**Red areas have more new immigrants (people moving to the UK in the last 2 years) than 98% of Scotland.**

```{r}
#new immigrants
aberdeen_priority_sf %>%
  filter(new_to_uk_rank_scot <= 2) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}
aberdeen_priority %>%
  mutate(new_to_uk_perc = new_to_uk_prop * 100) %>%
  select(data_zone, dz_name, total_population, new_to_uk_perc) %>%
  arrange(desc(new_to_uk_perc)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by most new immigrants to UK") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    new_to_uk_perc = "% immigrated in last 2 years")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Row
----------------------------------------------

### 

**Red areas have more people of minority ethnic backgrounds than 97% of Scotland.**

```{r}
#minority ethnic
aberdeen_priority_sf %>%
  filter(minority_ethnic_rank_scot <= 5) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}
aberdeen_priority %>%
  mutate(minority_ethnic_perc = minority_ethnic_prop * 100) %>%
  select(data_zone, dz_name, total_population, minority_ethnic_perc) %>%
  arrange(desc(minority_ethnic_perc)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by most people of minority ethnic backgrounds") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    minority_ethnic_perc = "% minority ethnic background")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Health
==============================================

Row
----------------------------------------------

### 

Here are **two measures of health in Priority Neighbourhoods**

-   Where do people have more emergency hospitalisations?

-   Where do people have more hospitalisations for drug use?

Row
----------------------------------------------

### 

**Red areas have more emergency hospitalisations than 95% of Scotland.**

```{r}
#emergency care
aberdeen_priority_sf %>%
  filter(emergency_rank_scot <= 5) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}
aberdeen_priority %>%
  select(data_zone, dz_name, total_population, emerg) %>%
  arrange(desc(emerg)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by most emergency hospitalisations") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    emerg = "emergency hospitalisations (Scottish average = 100)")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Row
----------------------------------------------

### 

**Red areas have more hospitalisations for drug use than 95% of Scotland.**

```{r}
#mortality
aberdeen_priority_sf %>%
  filter(drug_rank_scot <= 5) %>%
  mapview(
    map.types = "OpenStreetMap",
    zcol = "name",
    legend = FALSE,
    alpha.regions = 0, color = "red", lwd = 2)
```

### 

```{r}

aberdeen_priority %>%
  select(data_zone, dz_name, total_population, drug) %>%
  arrange(desc(drug)) %>%
  gt() %>%
  tab_header(
    title = "Priority Neighbourhoods", 
    subtitle = "ordered by highest drug hospitalisations") %>%
  cols_label(
    data_zone = "Data Zone",
    dz_name = "Name", 
    total_population = "Population", 
    drug = "drug hospitalisations (Scottish average = 100)")  %>%
  tab_options(container.height = 400,
              data_row.padding = 1,
              heading.padding = 1,
              column_labels.padding = 1)
```

Sources
==============================================

Row
----------------------------------------------
### 
Questions and comments very welcome!  
Email [Jess Butler](jessicabutler@abdn.ac.uk)  


All data and code for this analysis are available [here](https://github.com/JessButler/priority_neighbourhoods)  

[Scottish Index of Multiple Deprivation 2020 version 2 ranks](https://www.gov.scot/publications/scottish-index-of-multiple-deprivation-2020v2-ranks/) were the source for  

- area ID (Data_Zone) 
- council area 
- total population 
- overall disadvantage score (percentile for SIMD2020v2_rank)

[Scottish Index of Multiple Deprivation 2020 version 2 indicators](https://www.gov.scot/collections/scottish-index-of-multiple-deprivation-2020/) were the source for  

- overcrowding 
- no central heating 
- people age 16 to 19 not in education, employment or training

[2011 Scottish census](https://www.scotlandscensus.gov.uk/documents/2011-census-table-data-sns-data-zone-2011/) was the source for  

- ethnicity (KS201SC.csv)
- recent immigration (QS803SC.csv)
- lone parents (QS112SC.csv)
- children under age 5 (LC1117SC.csv)

Data Zone map boundaries were from the [2011 Scotland Data Zones shapefiles](https://data.gov.uk/dataset/ab9f1f20-3b7f-4efa-9bd2-239acf63b540/data-zone-boundaries-2011)
